Supplementary Note: Trajectory Inference Wrappers

### 2018-03-23 analysis/paper/supp_notes/supp_note_trajectory_inference_methods.Rmd
move aga supp note information

# load libraries
library(dynalysis)
library(dynplot)
library(tidyverse)
library(cowplot)
```

# `aga`: AGA
<insert description AGA>

## Parameters

```{r}
method <- description_aga()
par_set <- method$par_set
par_set
```

The default parameters are the following:
```{r}
defprm <- ParamHelpers::dfRowToList(
  ParamHelpers::generateDesignOfDefaults(par_set, trafo = TRUE),
  par_set, 
  i = 1
)
defprm
```

## Example dataset
We will be using a consecutive bifurcating toy dataset to demonstrate this method with.

```{r}
toy <- dyntoy::toy_tasks %>% filter(model == "consecutive_bifurcating") %>% extract_row_to_list(row_id = 1)
expression <- toy$expression

@@

Since AGA is implemented in Python, we created an R package for it. 
```{r}
aga_out <- aga::aga(
  counts = expression,
  start_cell = NULL,
  grouping_assignment = NULL,
  n_neighbours = defprm$n_neighbours,
  n_pcs = defprm$n_pcs,
  n_dcs = defprm$n_dcs,
  resolution = defprm$resolution,
  tree_based_confidence = defprm$tree_based_confidence,
  verbose = FALSE,
  num_cores = 1
)
```

<Talk about aga_out$adj>
```{r}
aga_out$adj %>% head(6) %>% knitr::kable()
```

<Talk about aga_out$obs>
```{r}
aga_out$obs %>% head(6) %>% knitr::kable()
```

Reformat everything into milestone_network and milestone_percentages
```{r}
cell_ids <- rownames(expression)

milestone_assignment_cells <- setNames(aga_out$obs$group_id, aga_out$obs$cell_id)

milestone_network <- aga_out$adj %>%
  mutate_at(vars(from, to), as.character) %>%
  filter(aga_adjacency_tree_confidence > 0) %>%
  mutate(
    length = 1,
    directed = TRUE
  ) %>%
  select(from, to, length, directed)

milestone_ids <- sort(unique(milestone_assignment_cells))
```

The prediction is wrapped using the `add_trajectory_to_wrapper` function. 
```{r}
prediction <- wrap_prediction_model(
  cell_ids = rownames(expression)
) %>% add_cluster_graph_to_wrapper(
  milestone_ids = milestone_ids,
  milestone_network = milestone_network,
  milestone_assignment_cells = milestone_assignment_cells,
  aga_out = aga_out
)

# `atan`: Arc-tangent
This is a control method for inferring cyclic trajectories in a very simple approach,
consisting of a dimensionality reduction step and followed by calculating the arc-tangent
between the first and the second component.

## Parameters
There is only one parameter, namely the dimensionality reduction method used.
```{r}
method <- description_atan()
par_set <- method$par_set
par_set
```

The default parameters are the following:
```{r}
defprm <- ParamHelpers::dfRowToList(
  ParamHelpers::generateDesignOfDefaults(par_set, trafo = TRUE),
  par_set, 
  i = 1
)
defprm
```

@@

```{r}
dimred <- dynmethods:::dimred(expression, method = defprm$dimred, ndim = 2)

>>>>

source(paste0(dynalysis::get_dynalysis_folder(), "/analysis/paper/setup.R"))
```

This document serves to provide insight into how each trajectory inference method was used and wrapped.

```{r}
library(dynmethods)
library(dynwrap)
library(dynutils)
library(dynplot)
```

<!-- ################################################## aga ################################################## -->
See [https://github.com/dynverse/dynmethods/blob/master/vignettes/aga.Rmd](dynmethods).

<!-- ################################################## agapt ################################################## -->

# `agapt`: AGA pseudotime
TODO: move to dynmethods.

When a start cell is given, AGA supports ordering the cells along branches.

## Parameters
```{r}
method <- description_agapt()
method$par_set
```

## Example dataset
We will be using the same consecutive bifurcating toy dataset.

```{r}
toy <- dyntoy::toy_tasks %>% 
  filter(model == "consecutive_bifurcating") %>% 
  extract_row_to_list(row_id = 1)
expression <- toy$expression
start_cells <- toy$prior_information$start_cells

@@

Since AGA is implemented in Python, we created an R package for it, available at [github.com/dynverse/aga](https://github.com/dynverse/aga).
The R package will write the expression values and any parameters to a file, execute a python script, and read the results from output files produced by AGA.
```{r}
defprm <- get_default_parameters(method)

# sample one start cell
start_cell <- sample(start_cells, 1)

aga_out <- aga::aga(
  expression = expression,
  start_cell = start_cell,
  grouping_assignment = NULL,
  n_neighbours = defprm$n_neighbours,
  n_pcs = defprm$n_pcs,
  n_dcs = defprm$n_dcs,
  resolution = defprm$resolution,
  tree_based_confidence = defprm$tree_based_confidence,
  verbose = TRUE,
  num_cores = 1
)
```

After building a kNN graph (not shown), the kNN graph is clustered into louvain groups:
```{r}
aga_out$obs %>% head(6) %>% knitr::kable()

# cell_ids <- rownames(expression)
# milestone_assignment_cells <- setNames(aga_out$obs$group_id, aga_out$obs$cell_id)
```

Several tests are used to assess which transitions exist between the louvain groups. 
We use the `aga_adjacency_tree_confidence` to construct the milestone network.
```{r}
aga_out$adj %>% head(6) %>% knitr::kable()
```

<!-- ################################################## atan ################################################## -->

# `atan`: Arc-tangent
TODO: move to dynmethods.

This is a control method for inferring cyclic trajectories in a very simple approach,
consisting of a dimensionality reduction step and followed by calculating the arc-tangent
between the first and the second component.

## Parameters
```{r}
method <- description_atan()
method$par_set
```

@@

```{r}
defprm <- get_default_parameters(method)

dimred <- dynmethods:::dimred(expression, method = defprm$dimred, ndim = 2)
